---
title: "Take-home Exercise 1: Investigating Student Performance with Data Visualisations"
author: "Alexei Jason"
date: "January 19, 2024"
date-modified: "January 21, 2024"
execute: 
  eval: true
  echo: true
  warning: false
editor: visual
---

# 1 Overview

The OECD Programme for International Student Assessment (PISA) did a deep dive into how 15-year-old students are doing in reading, math, and science. They didn't just stop there – they also dug into students' backgrounds, home situations, and school factors that could shake things up. PISA is like the gold standard when it comes to checking out how well students are really learning on a global scale. The results spill the beans on the quality and fairness of learning achievements worldwide, giving educators and policy folks the lowdown on what's working in different countries.

But PISA isn't just about the numbers; it's about creating a common ground and getting people to team up and use that data. The whole idea is to get schools and policymakers to stop sizing themselves up against their own kind and start looking outwards. PISA wants them to peek at what teachers, schools, and policymakers are doing around the world and learn a thing or two.

Fast-forward to 2022, and PISA is still at it. They've got eight questionnaires floating around, getting the scoop from different education big shots – students, teachers, parents, and school principals. It's like a global report card but with a whole lot more details.

# 2 Data Preparation

## 2.1 Importing packages

In this take home exercise we are going to utilise several package

1.  Tidyverse (to wrangle and plot our data)
2.  Haven (to read our dataset from sas data type)
3.  Patchwork (to plot multiple plot in the same figure)

```{r}
pacman::p_load(tidyverse, haven, patchwork)
```

## 2.2 Importing dataset

Importing the data in sas data type

```{r}
#| eval: false
stu_qqq <- read_sas('data/cy08msp_stu_qqq.sas7bdat')
```

Filtering the data to singapore

```{r}
#| eval: false
stu_qqq_SG <- stu_qqq %>%
  filter(CNT == "SGP")
```

Export the filtered dataset to rds to ease further data importing

```{r}
#| eval: false
write_rds(stu_qqq_SG, 'data/stu_qqq_SG.rds')
```

Read the exported rds

```{r}
stu_qqq_SG <- read_rds('data/stu_qqq_SG.rds')
```

## 2.3 Variable Preparations

In preparing for our analysis, we'll filter out the columns we need and focus on six variables:

1.  **`school_type`** (Renamed from STRATUM)

2.  **`gender`** (Renamed from ST004D01T)

3.  **`math`** (Average of 10 math scores)

4.  **`science`** (Average of 10 science scores)

5.  **`reading`** (Average of 10 reading scores, renamed from 'read')

6.  **`social_economic`** (Renamed from ESCS)

Additionally, we'll recode the answers of questions to more descriptive measures and adjust the data types of the columns for clarity and consistency.

For example, if the gender column uses numerical codes, we can replace them with descriptive labels (e.g., 1 for male, 2 for female can be recoded as 'Male' and 'Female').

Lastly, we'll ensure that the data types of each column align with the nature of the information they hold (e.g., converting numerical columns to appropriate numeric types).

```{r}
stu_filtered <- stu_qqq_SG %>%
  mutate(math = rowMeans(select(stu_qqq_SG,
         PV1MATH,PV2MATH,PV3MATH,PV4MATH,PV5MATH,PV6MATH,PV7MATH,PV8MATH,PV9MATH,PV10MATH)),
         science = rowMeans(select(stu_qqq_SG,
         PV1SCIE,PV2SCIE,PV3SCIE,PV4SCIE,PV5SCIE,PV6SCIE,PV7SCIE,PV8SCIE,PV9SCIE,PV10SCIE)),
         reading = rowMeans(select(stu_qqq_SG,
         PV1READ,PV2READ,PV3READ,PV4READ,PV5READ,PV6READ,PV7READ,PV8READ,PV9READ,PV10READ))
         ) %>%
  select(STRATUM,ST004D01T,math,science,reading,ESCS) %>%
  rename(c(school_type = "STRATUM", 
           gender = "ST004D01T",
           socio_economic = "ESCS")) %>%
  mutate(gender = recode(gender, 
                         "1" = "Female", 
                         "2" = "Male"))

stu_filtered$school_type <- as.factor(stu_filtered$school_type)
stu_filtered$gender <- as.factor(stu_filtered$gender)
```

## 2.4 Overview of the Data-set

Here is the overview of the data-set that we are going to use for the analysis:

```{r}
glimpse(stu_filtered)
```

## 2.5 Handling Missing Values

Now, we are going through our data and getting rid of rows that are missing some info. The main reason for doing this is because later on, we want to make our variables easier to understand. If there are blank spots in our data columns, it might mess up the way we simplify things if a value accidentally ends up in a spot that was supposed to be blank.

We use the `is.na()` to spot those gaps in our data, and then `sum()` helps us count how many times these missing values pop up. It's all part of making our data more user-friendly.

```{r}
sum(is.na(stu_filtered$school_ID))
```

```{r}
sum(is.na(stu_filtered$gender))
```

```{r}
sum(is.na(stu_filtered$math))
```

```{r}
sum(is.na(stu_filtered$science))
```

```{r}
sum(is.na(stu_filtered$reading))
```

```{r}
sum(is.na(stu_filtered$socio_economic))
```

The variable with missing values is: *socio_economic.*

Now, we are going to drop variable which contains missing values using **'drop_na()'**

```{r}
stu_filtered <- drop_na(stu_filtered,socio_economic)

stu_filtered
```

The output shows that 6,559 observations are retained.

# 3 Exploratory Data Analysis (EDA)

In this section 4 analysis will be performed to get insights and uncover truth of factors for Singapore student performances.

### 3.1 Student Performances

```{r}
stu_grade <- stu_filtered %>%
  select (math, science, reading)
```

```{r}
#| code-fold: true
#| code-summary: "Show the code"
# Distribution of Performance in Mathematics
P1 <- ggplot(data = stu_grade,
       aes(x = math)) +
  geom_density(color = "#459395", size = 0.6, fill= "#459395", alpha = 0.4) +
  coord_cartesian(xlim = c(0,1000)) +
  geom_vline(aes(xintercept = mean(math)),
             color = "red", alpha = 0.8, linewidth = 0.7, linetype = "dashed") +
  annotate("text", x = 400, y = 0.0035,
           label = paste("Mean=", 
                         round(mean(stu_grade$math, na.rm=T), 2)),
           color = "red", size = 3) +
  geom_vline(aes(xintercept = median(math)),
             color= "grey50", linewidth = 0.7, linetype = "solid") +
  annotate("text", x = 800, y = 0.0035,
           label = paste("Median=", 
                         round(median(stu_grade$math, na.rm=T), 2)),
           color = "grey20", size = 3) +  
  geom_boxplot(width = 0.0005, fill = "white", alpha = 0.5,
               position = position_nudge(y = -0.0005)) +
  theme_minimal()+
  labs(title="Distribution of Performance in Mathematics") +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title=element_text(size= 12),
        axis.text = element_text(size= 8)) 

# Distribution of Performance in Reading
P2 <- ggplot(data = stu_grade,
       aes(x = reading)) +
  geom_density(color = "#EB7C69", size = 0.6, fill= "#EB7C69", alpha = 0.4) +
  coord_cartesian(xlim = c(0,1000)) +
  geom_vline(aes(xintercept = mean(reading)),
             color = "red", alpha = 0.8, linewidth = 0.7, linetype = "dashed") +
  annotate("text", x = 400, y = 0.0035,
           label = paste("Mean=", 
                         round(mean(stu_grade$reading, na.rm=T), 2)),
           color = "red", size = 3) +
  geom_vline(aes(xintercept = median(reading)),
             color= "grey50", linewidth = 0.7, linetype = "solid") +
  annotate("text", x = 800, y = 0.0035,
           label = paste("Median=", 
                         round(median(stu_grade$reading, na.rm=T), 2)),
           color = "grey20", size = 3) +  
  geom_boxplot(width = 0.0005, fill = "white", alpha = 0.5,
               position = position_nudge(y = -0.0005)) +
  theme_minimal()+
  labs(title="Distribution of Performance in Reading",
       y = "density") +
  theme(axis.title.x = element_blank(),
        plot.title=element_text(size= 12),
        axis.text = element_text(size= 8)) 
  
# Distribution of Performance in Science
P3 <- ggplot(data = stu_grade,
       aes(x = science)) +
  geom_density(color = "#FDA638", size = 0.6, fill= "#FDA638", alpha = 0.4) +
  coord_cartesian(xlim = c(0,1000)) +
  geom_vline(aes(xintercept = mean(science)),
             color = "red", alpha = 0.8, linewidth = 0.7, linetype = "dashed") +
  annotate("text", x = 400, y = 0.0035,
           label = paste("Mean=", 
                         round(mean(stu_grade$science, na.rm=T), 2)),
           color = "red", size = 3) +
  geom_vline(aes(xintercept = median(science)),
             color= "grey50", linewidth = 0.7, linetype = "solid") +
  annotate("text", x = 800, y = 0.0035,
           label = paste("Median=", 
                         round(median(stu_grade$science, na.rm=T), 2)),
           color = "grey20", size = 3) +  
  geom_boxplot(width = 0.0005, fill = "white", alpha = 0.5,
               position = position_nudge(y = -0.0005)) +
  theme_minimal()+
  labs(title="Distribution of Performance in Science") +
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title=element_text(size= 12),
        axis.text = element_text(size= 8)) 


P1 / P2 / P3
```

Based on the provided figure, it appears that the performance distribution follows a normal pattern. However, there's a slight variation in performance levels, with math scores being higher than science, and science scores being higher than reading.

Moreover, it's noticeable that the performance variance is quite substantial. This suggests that students in Singapore exhibit significant differences in their academic achievements. Several factors, including school, gender, and socioeconomic status, could contribute to these variations in performance.

### 3.2 Gender

The first factor that we are going to look into is gender.

```{r}
ggplot(data = stu_filtered,
       aes(x = gender)) +
  geom_bar(width = 0.5,
           color = "black",
           fill = "#01796F")+
  theme_minimal()
```

From the distribution graph above can be seen that the questionnaire have similar number of male and female student.

::: panel-tabset
#### Math

Observing the graph below, it's evident that there are **minimal distinctions** between male and female math scores. Males exhibit a **slightly higher** **average and median score**, though the differences are quite small

```{r}
#| code-fold: true
#| code-summary: "Show the code"
stats <- data.frame(gender = as.factor(c('Female','Male')),
                    mean = c(mean(stu_filtered$math[stu_filtered$gender == "Female"]),
                             mean(stu_filtered$math[stu_filtered$gender == "Male"])),
                    median = c(median(stu_filtered$math[stu_filtered$gender == "Female"]),
                               median(stu_filtered$math[stu_filtered$gender == "Male"])) )

ggplot(data = stu_filtered,
       aes(y = math)) +
  geom_histogram(bins = 10,
                 color = 'grey50',
                 fill = 'grey90') +
  facet_wrap(~ gender) +
  geom_hline(data = stats,
             mapping = aes(yintercept=mean),
             color = "red") +
  geom_hline(data = stats,
             mapping = aes(yintercept=median),
             color = "blue") +
  annotate(
    'text',
    x = Inf,
    y = Inf,
    hjust = 1,
    vjust = 2.1,
    label = paste("mean"),
    color = "red"
  ) +
  annotate(
    'text',
    x = Inf,
    y = Inf,
    hjust = 1,
    vjust = 1,
    label = paste("median"),
    color = "blue"
  ) +
  theme_minimal()
```

#### Science

Observing the graph provided, it's evident that there are **minimal disparities** between male and female science scores. Males exhibit a **slightly higher average and median score**, though the differences are quite small.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
stats <- data.frame(gender = as.factor(c('Female','Male')),
                    mean = c(mean(stu_filtered$science[stu_filtered$gender == "Female"]),
                             mean(stu_filtered$science[stu_filtered$gender == "Male"])),
                    median = c(median(stu_filtered$science[stu_filtered$gender == "Female"]),
                               median(stu_filtered$science[stu_filtered$gender == "Male"])) )

ggplot(data = stu_filtered,
       aes(y = science)) +
  geom_histogram(bins = 10,
                 color = 'grey50',
                 fill = 'grey90') +
  facet_wrap(~ gender) +
  geom_hline(data = stats,
             mapping = aes(yintercept=mean),
             color = "red") +
  geom_hline(data = stats,
             mapping = aes(yintercept=median),
             color = "blue") +
  annotate(
    'text',
    x = Inf,
    y = Inf,
    hjust = 1,
    vjust = 2.1,
    label = paste("mean"),
    color = "red"
  ) +
  annotate(
    'text',
    x = Inf,
    y = Inf,
    hjust = 1,
    vjust = 1,
    label = paste("median"),
    color = "blue"
  ) +
  theme_minimal()
```

#### Reading

Based on the graph below, it's noticeable that there are very slight distinctions between male and female reading scores. Males tend to have a **slightly lower average and median score**, though the differences are minimal.

```{r}
#| code-fold: true
#| code-summary: "Show the code"
stats <- data.frame(gender = as.factor(c('Female','Male')),
                    mean = c(mean(stu_filtered$reading[stu_filtered$gender == "Female"]),
                             mean(stu_filtered$reading[stu_filtered$gender == "Male"])),
                    median = c(median(stu_filtered$reading[stu_filtered$gender == "Female"]),
                               median(stu_filtered$reading[stu_filtered$gender == "Male"])) )
ggplot(data = stu_filtered,
       aes(y = reading)) +
  geom_histogram(bins = 10,
                 color = 'grey50',
                 fill = 'grey90') +
  facet_wrap(~ gender) +
  geom_hline(data = stats,
             mapping = aes(yintercept=mean),
             color = "red") +
  geom_hline(data = stats,
             mapping = aes(yintercept=median),
             color = "blue") +
  annotate(
    'text',
    x = Inf,
    y = Inf,
    hjust = 1,
    vjust = 2.1,
    label = paste("mean"),
    color = "red"
  ) +
  annotate(
    'text',
    x = Inf,
    y = Inf,
    hjust = 1,
    vjust = 1,
    label = paste("median"),
    color = "blue"
  ) +
  theme_minimal()
```
:::

Overall, can be concluded that student's performances has **SLIGHLTY VARIATION** between both gender.

### 3.3 School

The student that was participating in this questionnaire was from 164 different school in Singapore

On the code above we are going to calculate the average of performance of student for each school across Singapore against Singapore overall average.

Followed by calculating the error or the differences between each school average and the average in Singapore.

```{r}
#| code-fold: true
#| code-summary: "Show the code"

P7 <- ggplot(data= stu_filtered,
       aes(x= school_type, y= math)) +
  geom_violin(color = "#459395", size = 0.6, fill= "#459395", alpha = 0.4) +
  geom_boxplot(width= 0.4, outlier.colour = "grey30", outlier.size = 2, 
               outlier.alpha = 0.5, outlier.shape = 19) +
  stat_summary(geom = "point",       
               fun.y="mean",         
               colour ="red",        
               size=3) +  coord_cartesian(ylim = c(0,1000)) +
  scale_color_manual(values=c("#999999", "#E69F00")) +
  theme_minimal() +
  labs(title="Mathematics") +  
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        plot.title=element_text(size= 12, hjust= 0.5),
        axis.text = element_text(size= 10)) 
  
P8 <- ggplot(data= stu_filtered,
       aes(x= school_type, y= reading)) +
  geom_violin(color = "#EB7C69", size = 0.6, fill= "#EB7C69", alpha = 0.4) +
  geom_boxplot(width= 0.4, outlier.colour = "grey30", outlier.size = 2, 
               outlier.alpha = 0.5, outlier.shape = 19) +
  stat_summary(geom = "point",       
               fun.y="mean",         
               colour ="red",        
               size=3) +  coord_cartesian(ylim = c(0,1000)) +  
  theme_minimal() +
  labs(title="Reading") + 
  theme(axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        plot.title=element_text(size= 12, hjust= 0.5),
        axis.text = element_text(size= 10)) 
  
P9 <- ggplot(data= stu_filtered,
       aes(x= school_type, y= science)) +
  geom_violin(color = "#FDA638", size = 0.6, fill= "#FDA638", alpha = 0.4) +
  geom_boxplot(width= 0.4, outlier.colour = "grey30", outlier.size = 2, 
               outlier.alpha = 0.5, outlier.shape = 19) +
  stat_summary(geom = "point",       
               fun.y="mean",         
               colour ="red",        
               size=3) +  coord_cartesian(ylim = c(0,1000)) +  
  theme_minimal() +
  labs(title="Science") + 
  theme(axis.title.x = element_blank(),
        axis.text.y = element_blank(),
        axis.title.y = element_blank(),
        plot.title=element_text(size= 12, hjust= 0.5),
        axis.text = element_text(size= 10)) 

(P7 + P8 + P9) +
    plot_annotation(title= "School-Based Performance Comparison ",
                    theme = theme(plot.title=element_text(size= 15, hjust= 0.5)))
```

The plots above show that the differences between all study subject were similar. However, the differences between a school and the Singapore average are ranged from 0 to 220. And, the average differences is around 40. This mean that not every single school in Singapore are having the same performance with each other.

On the other hand, the q3 of the box plot lay around 70 point scores. This mean that almost 75% of school in Singapore are located between 0-70 point difference to national average. On the other side, there are some school that are having almost 150 point difference either higher or lower against the national average.

### 3.4 Social Economic

The last factor that are going to be explored is social economic. The social economic range from -3.55 to 3.28.

```{r}

summary(stu_filtered$socio_economic)
```

The code below plot the distribution of student social economic standing in Singapore.

```{r}
#| code-fold: true
#| code-summary: "Show the code"

ggplot(data = stu_filtered,
       aes(y = socio_economic)) +
  geom_violin(aes(x = 0))+
  geom_boxplot()+
  stat_summary(geom = 'point',
               fun = 'mean',
               aes(x = 0),
               color = 'red') +
  scale_y_continuous(NULL) +
  scale_x_continuous('Social Economic Standing', breaks = NULL) +
  theme_minimal()
```

```{r}
#| code-fold: true
#| code-summary: "Show the code"
cor1 <- round(cor(stu_filtered$math, stu_filtered$socio_economic),2)
cor2 <- round(cor(stu_filtered$science, stu_filtered$socio_economic),2)
cor3 <- round(cor(stu_filtered$reading, stu_filtered$socio_economic),2)

p1 <- ggplot(data = stu_filtered,
             aes(y = math, x = socio_economic)) +
        geom_point(size = 0.1)+
        geom_smooth(method = lm) + 
        annotate("text", x = 2.5, y = 600, label=paste0("r = ", cor1), color = 'blue') +
        theme_minimal()

p2 <- ggplot(data = stu_filtered,
             aes(y = science, x = socio_economic)) +
        geom_point(size = 0.1)+
        geom_smooth(method = lm) +
        annotate("text", x = 2.5, y = 600, label=paste0("r = ", cor2), color = 'blue') + 
        theme_minimal()

p3 <- ggplot(data = stu_filtered,
             aes(y = reading, x = socio_economic)) +
        geom_point(size = 0.1)+
        geom_smooth(method = lm) + 
        annotate("text", x = 2.5, y = 600, label=paste0("r = ", cor3), color = 'blue') +
        theme_minimal()

p1/p2/p3
```

From the three plots, it is apparent that the scatter is broadly distributed across all levels. However, when examining the correlation between social-economic status and academic performances, it is consistently around 0.44. This indicates a weak to medium positive correlation between social-economic standing and students' performances. In other words, it can be inferred that students with higher social-economic standing tend to achieve higher scores.

# 4 Conclusion

These observations suggest several key points about the educational landscape in Singapore:

1.  **Similarity Across Subjects**: The consistency in student performances across different subjects implies that factors affecting academic achievement may be relatively consistent across disciplines. This could include teaching quality, curriculum standards, and student support mechanisms.

2.  **Gender Neutrality in Performance**: The lack of a significant impact of gender on student performance suggests that academic success in Singapore is not heavily influenced by gender-specific factors. This could reflect a culture of equality and fairness in education.

3.  **Variations Among Schools**: The variations in performance levels among schools highlight disparities in educational resources, teaching methodologies, and perhaps even socio-cultural environments. This suggests that efforts to standardize and equalize educational opportunities across schools may be necessary.

4.  **Minor Role of Socio-economic Standing**: The relatively minor influence of socio-economic standing on academic performance indicates that Singapore's education system may effectively mitigate the effects of socio-economic disparities. This could be attributed to policies aimed at providing equal access to education and support for students from diverse socio-economic backgrounds.

Overall, these observations underscore the complex interplay of various factors shaping educational outcomes in Singapore, highlighting the need for continued efforts to address disparities and ensure equitable access to quality education for all students.
